---
title: "Merging physiology and symbiont dataframes"
output: html_document
date: "2024-06-14"
editor_options: 
  chunk_output_type: console
---

# Set up 

```{r}
library(tidyverse)
library(data.table)
```



# Read in ITS2 data and summarize by colony_id

Read in ITS2 data and summarize by colony_id, because symbiont communities didnt change over time. Calculate this as the average relative abundance of each taxa for each colony. 

## Read in data 

```{r}
sym<-fread("data/coral_physiology/symbiont_rel_abund.csv")
```

## Calculate mean abundance for each taxa by colony ID

```{r}
result <- sym[, lapply(.SD, mean, na.rm = TRUE), by = .(colony_id, site, host_genus), .SDcols = sapply(sym, is.numeric)]
```

## Calculate the sum of all taxa to check that relative abundance does not exceed 1 

```{r}
# Calculate the sum of all numeric columns for each row and add it as a new column
sums_result <- result[, total_sum := rowSums(.SD, na.rm = TRUE), .SDcols = sapply(result, is.numeric)]

# Calculate the sum of the new column grouped by colony_id
sums <- result[, .(total_sum = sum(total_sum)), by = colony_id]
```

All values add to 1. 

# Merge files together - symbiont composition and physiology

## Read in data files 

```{r}
phys<-read_csv("data/coral_physiology/timeseries_physiology.csv")
```

## Merge files by colony_id

Merge using left_join, such that all observations in phys are kept 

```{r}
merged <- full_join(phys, result)
merged <- merged %>%
  select(!total_sum)
```

## Write new merged file to output 

```{r}
write_csv(merged, file="data/coral_physiology/merged_data.csv")
```

